```{r front, child="front.Rmd", echo=FALSE}
```

## Dependencies

This document has the following dependencies:

```{r dependencies, warning=FALSE, message=FALSE}
library(GenomicRanges)
library(rtracklayer)
library(AnnotationHub)
library(TFutils)
```

## Use TFutils to get ENCODE 690 dataset

```{r tfutils}
names(TFutils::encode690)
TFutils::encode690[,1:5]
df1 = TFutils::encode690["title"][1]
df1[[1]]
```

Lets take a look at one dataset for the TF for the BRCA1 gene from the Gm12878 cell line:

```{r ahub_GR}
ah = AnnotationHub()
tfbs_title = df1[[1]][286] # wgEncodeAwgTfbsBroadHuvecCtcfUniPk.narrowPeak.gz
gr1 <- subset(ah, title == tfbs_title)[[1]]
gr1
```

There are 551 TFs in the dataset. Let's look at the distribution of sizes.

```{r ahub_summary}
summary(width(gr1))
table(width(gr1))
```

Most peaks in gr1 have a width of 296bp.

Let's get the center 101bp coordinates of the TF.

```{r ahub_resize}
dnabert_input = resize(gr1, width = 101, fix = "center")
table(width(dnabert_input))
```

Look at sequences

```{r get_sequences}
library(memes)
human_genome = BSgenome.Hsapiens.UCSC.hg19
dnabert_input_dss = get_sequence(dnabert_input, human_genome)
dnabert_input_dss
```

Convert to data frame; generate negative sequences by shuffling

```{r reformat_and_shuffle_for_negative}
library(universalmotif)

dss2df = function(dss) data.frame(names=names(dss), seq=as.character(dss), neg_seq=shuffle_sequences(dss, k=2, method="euler"))

dnabert_input_df = dss2df(dnabert_input_dss)
dnabert_input_df
```

Save in a csv file

```{r save_csv}
library(stringr)
dsname_long = tfbs_title
dsname_pre = str_replace_all(dsname_long,".narrowPeak.gz","")
dsname = str_replace_all(dsname_pre,"wgEncodeAwgTfbs","")
filename=paste("/Users/sjg/Downloads/",dsname,".csv", sep='')
write.csv(dnabert_input_df, file=filename)
```

Function to create positive and negative sequences from each TFBS dataset

```{r ahub_seq}
ds_create = function(ah, ds) {
  gr = subset(ah, title == ds)[[1]]
  dnabert_input = resize(gr, width = 101, fix = "center")
  human_genome = BSgenome.Hsapiens.UCSC.hg19
  dnabert_input_dss = get_sequence(dnabert_input, human_genome)
  dnabert_input_df = dss2df(dnabert_input_dss)
  dsname_long = ds
  dsname_pre = str_replace_all(dsname_long,".narrowPeak.gz","")
  dsname = str_replace_all(dsname_pre,"wgEncodeAwgTfbs","")
  filename=paste("/Users/sjg/Downloads/tfbs_datasets/",dsname,".csv", sep='')
  write.csv(dnabert_input_df, file=filename)
}
```

Loop over first 10 tfbs datasets

```{r loop_tfbs}
for (tfbs_ds in df1[[1]][101:140]) {
  ds_create(ah,tfbs_ds)
}
```

## SessionInfo

\scriptsize

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
