```{r front, child="front.Rmd", echo=FALSE}
```

## Dependencies

This document has the following dependencies:

```{r dependencies, warning=FALSE, message=FALSE}
library(GenomicRanges)
library(rtracklayer)
library(AnnotationHub)
library(TFutils)
```

## Use TFutils to get ENCODE 690 dataset

```{r tfutils}
names(TFutils::encode690)
TFutils::encode690[,1:5]
df1 = TFutils::encode690["title"][1]
df1[[1]]
```

Lets take a look at one dataset for the TF for the BRCA1 gene from the Gm12878 cell line:

```{r ahub_GR}
ah = AnnotationHub()
tfbs_title = df1[[1]][1] # wgEncodeAwgTfbsBroadHuvecPol2bUniPk.narrowPeak.gz
gr1 <- subset(ah, title == tfbs_title)[[1]]
gr1
```

There are 551 TFs in the dataset. Let's look at the distribution of sizes.

```{r ahub_summary}
summary(width(gr1))
table(width(gr1))
```

Most peaks in gr1 have a width of 296bp.

Let's get the center 101bp coordinates of the TF.

```{r ahub_resize}
dnabert_input = resize(gr1, width = 101, fix = "center")
table(width(dnabert_input))
```

Look at sequences

```{r get_sequences}
library(memes)
human_genome = BSgenome.Hsapiens.UCSC.hg19
dnabert_input_dss = get_sequence(dnabert_input, human_genome)
dnabert_input_dss
```

```{r file_test}
df1[[1]][661:690]
```


Find random sequence with width of 101 and same GC content

```{r helper_functions}
getichrpos = function(ipos){
      j = order(ipos); 
      ipos = sort(ipos); 
      ci = 1;
      res = rep(NA, length(ipos))
      for(i in 1:length(ipos))
      {
        while(ipos[i]>chrpos[ci]){
          ci = ci+1; 
        }
        res[j[i]] = ci
      }
      return(res); 
}

# generateRandomGenSeq 
#
# input: number of sequences
# output: GenomicRange of sequences
#
    
generateRandomGenSeqs = function(seqlens){
      rpos = sample(pmax, length(seqlens), replace = TRUE)
      ichr1 = getichrpos(rpos)
      ichr2 = getichrpos(rpos+seqlens)
      
      jj = which(ichr1!=ichr2)
      while(length(jj)>0){
        rpos[jj] = sample(pmax, length(jj), replace = TRUE)
        ichr1 = getichrpos(rpos)
        ichr2 = getichrpos(rpos+seqlens)
        jj = which(ichr1!=ichr2)
      }
      chr = ichrA[ichr1]
      start = rpos - chrpos0[ichr1];
      names <- chr; 
      ranges <- IRanges::IRanges(start=start, width=seqlens)
      strand <- BiocGenerics::strand(sample(c("+", "-"), length(names), replace=TRUE))
      gr <- GenomicRanges::GRanges(seqnames=names, ranges=ranges, strand=strand)
}

gcContent <- function(seqs)
    {
      alf <- Biostrings::alphabetFrequency(seqs, as.prob=TRUE)
      gc = rowSums(alf[,c("G", "C"), drop=FALSE])
}

## 

matchSeqs = function(gc1, gc2, len1, len2,  gc_th=0.02, len_th=0.02){
      ##
      #  gc1 = rndGC; 
      #  gc2 = inGC; 
      #  len1 = seqlens; 
      #  len2 = seqlens;
      #  rpt1 = inRpt; 
      #  rpt2 = rndRpt; 
      #  gc1=desGC[unmatched];gc2= rndGC;len1= desLens[unmatched];len2= BiocGenerics::width(rndBed);rpt1= desRpt[unmatched];rpt2= rndRpt;
                      
      #  gc_th=0.02
      #  len_th=0.02
      
      
      # gc1, len1 are the desired 
      # gc2 and len2 are to be matched 
      len_th = len_th * len1; 
      
      i1 = order(gc1)
      i2 = order(gc2)
      
      gc1 = gc1[i1]
      gc2 = gc2[i2]
      len1 = len1[i1]
      len2 = len2[i2]
      
      gc2 = c(gc2, 1E10)
      
      len_th = len_th[i1]
      
      m2 = 1; 
      N = length(i1); 
      N2 = length(i2);
      mtc1 = rep(NA, N)
      mtc2 = rep(0, length(i2))
      for(i in 1:N){
        #if(i%%1000==0){cat(i,' ')}
        gc1i = gc1[i]; 
        len1i = len1[i]
        len_thi = len_th[i]
        
        while(gc1i - gc2[m2]>gc_th) {
          m2 = m2+1; 
        }
        if(m2<=N2){
          m2b=m2;
          while(gc2[m2b]-gc1i<=gc_th){
            if ((mtc2[m2b]==0)&(abs(len1i-len2[m2b])<=len_thi)){
              mtc2[m2b]=i; 
              mtc1[i]=m2b;
              if(m2b==m2){m2 = m2+1;}
              break; 
            }
            m2b =m2b+1; 
          }
        }else{break;}
      }
      
      mtc1 = i2[mtc1]
      res = rep(NA, N)
      res[i1] = mtc1; 
      return(res)
}

```

```{r import_bed_file}

ds_create_with_random_neg = function(dnabert_input, ds, genome) {
  
  # create BED file
  
  df <- data.frame(seqnames=seqnames(dnabert_input),
  starts=start(dnabert_input)-1,
  ends=end(dnabert_input),
  names=c(rep(".", length(dnabert_input))),
  scores=c(rep(".", length(dnabert_input))),
  strands=strand(dnabert_input))
  write.table(df, file="dnabert_input.bed", quote=F, sep="\t", row.names=F, col.names=F)
  
  # set genome parameters
  
  seqnams = GenomeInfoDb::seqnames(human_genome)
  chrlens = GenomeInfoDb::seqlengths(human_genome)
  chrpos = cumsum(as.numeric(chrlens))
  pmax = max(chrpos)
  chrpos = c(chrpos,1E12)
  chrpos0 = c(0, chrpos)
  ichrA = as.character(names(chrlens))
  
  # import BED file representing dnabert input

  inBed = rtracklayer::import.bed(inputBedFN)
  inbed = GenomicRanges::as.data.frame(inBed)
  
  # extract sequences
  
  inSeqs = Biostrings::getSeq(human_genome, inBed)
  seqlens = inbed$width
  inGC = gcContent(inSeqs)
  
  # set up output matrix
  
  nout = round(nrow(inbed))
  outbed=matrix(ncol=ncol(inbed), nrow=nout)
  outSeq = rep(inSeqs, length=nout)
  colnames(outbed)=colnames(inbed)
  unmatched = 1:length(outSeq)
  desGC = rep(inGC, length=nout); #desired output GC
  desLens = rep(seqlens, length=nout); #desired output lengths 
  
  # find negative sequences
  
  nMaxTrials = 250
  batchsize = 7000
  length_match_tol = 0
  GC_match_tol = 0.01
  num_remain = 0
  for(iter in 1:nMaxTrials){
      if(length(unmatched)>0){
        cat(' Trial',iter,'out of',nMaxTrials,'\n')
        rndBed = generateRandomGenSeqs(rep(desLens[unmatched],length.out=batchsize))
        rndbed= GenomicRanges::as.data.frame(rndBed)
        cat(' importing sequences\n')
        rndSeqs = Biostrings::getSeq(genome, rndBed)
        rndGC = gcContent(rndSeqs)
        cat(' matching sequences\n')
        
        mtc = matchSeqs(desGC[unmatched], rndGC, desLens[unmatched], BiocGenerics::width(rndBed),
                        gc_th = GC_match_tol,
                        len_th = length_match_tol)
        jj = which(!is.na(mtc))
        if(length(jj)>0){
          #outbed[unmatched[jj],]=rndbed[mtc[jj],];
          outbed[unmatched[jj],1:5]=as.matrix(rndbed[mtc[jj],]);
          outSeq[unmatched[jj],]=rndSeqs[mtc[jj],];
          unmatched = unmatched[-jj]
        }
        cat(nrow(outbed) - length(unmatched),'sequences found so far, ',length(unmatched), ' remaining.\n')
      }
    if(length(unmatched) <= 2) {
      num_remain = num_remain + 1
      if (num_remain > 10) {
        break
      }
    }
  }  

  if(length(unmatched)>0){
    outbed = outbed[-unmatched,]
    outSeq = outSeq[-unmatched,]
    # added
    inbed = inbed[-unmatched,]
    inSeqs = inSeqs[-unmatched,]
  }
  
  # rearrange output
  
  outbed = gsub(' ','', outbed)
  #inseqnams = paste(as.character(inbed[,1]), ':', inbed[,2], '-', inbed[,3], 1:nrow(inbed), sep='')
  inseqnams = paste(as.character(inbed[,1]), ':', inbed[,2], '-', inbed[,3], sep='')
  #outseqnams = paste(outbed[,1], ':', outbed[,2], '-', outbed[,3], 1:nrow(outbed), sep='')
  outseqnams = paste(outbed[,1], ':', outbed[,2], '-', outbed[,3], sep='')
  #out_df = data.frame(pos_names=inseqnams, seq=as.character(inSeqs), neg_names=outseqnams, neg_seq = as.character(outSeq))
  out_df = data.frame(names=inseqnams, seq=as.character(inSeqs), neg_seq = as.character(outSeq))
  row.names(out_df) = make.names(inseqnams, unique=TRUE)
  
  # save to CSV file
  
  dsname_long = ds
  dsname_pre = str_replace_all(dsname_long,".narrowPeak.gz","")
  dsname = str_replace_all(dsname_pre,"wgEncodeAwgTfbs","")
  filename=paste("/Users/sjg/Downloads/tfbs_datasets/",dsname,"-ran.csv", sep='')
  write.csv(out_df, file=filename)
  
}


```

```{r test_function}
ds_create = function(ah, ds) {
  genome = BSgenome.Hsapiens.UCSC.hg19
  gr = subset(ah, title == ds)[[1]]
  dnabert_input = resize(gr, width = 101, fix = "center")
  ds_create_with_random_neg(dnabert_input, ds, genome)
}

```
Function to create positive and negative sequences from each TFBS dataset


Loop over first 10 tfbs datasets

```{r loop_tfbs}
for (tfbs_ds in df1[[1]][641:690]) {
  ds_create(ah,tfbs_ds)
}
```

```{r file_test}
df1[[1]][34:40]
```


```{r test_gc}
dna_pos = DNAString("CTGGCTCCCCCTGGCGGTCATGATAGGAAGCGCCCTTCCCCTCTGGCTCCTGTAGAGGGCGTCGCGGATGCATTGATCCTTTAGTTCCTTCACTGTCCATC")
dna_neg = DNAString("TTGACCGAAGCAGGGTGAGTGCCTTCACCAGTCGGTACCGGACGAGGCCGGGCTCCGGGCTAAGGGGCTGTCGAAACTGCTGAATAGGCGGAGAGGAAAAT")
dna_pos_GC = letterFrequency(dna_pos, "GC", as.prob = TRUE)
dna_neg_GC = letterFrequency(dna_pos, "GC", as.prob = TRUE)
dna_pos_GC
```

```{r test_overlap}
gr_ran = GRanges(seqnames=c("chr18", "chr7", "chr2", "chr2", "chr19", "chr9", "chr6", "chr13"),
                 strand="*",
                 ranges=IRanges(start = c(47370376,48075155,33864111,239793543,1567533,130916607,125283490,109801981),
                                  width = 101))
countOverlaps(dnabert_input, gr_ran)
```


## SessionInfo

\scriptsize

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
